---
title: "Using the smd() function"
author: "Bradley Saul"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Usage of smd}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
references:
- id: yang2012unified
  title: A unified approach to measuring the effect size between two groups using SAS
  author:
  - family: Yang
    given: Dongsheng
  - family: Dalton
    given: Jarrod E
  volume: 335
  URL: 'http://www.lerner.ccf.org/qhs/software/lib/stddiff.pdf'
  booktitle: SAS Global Forum
  page: 1--6
  type: article-journal
  issued:
    year: 2012
editor_options: 
  chunk_output_type: console
---


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


The `SugarMaryDenver` package provides the `smd` method to compute standardized mean differences between two groups for continuous values (`numeric` and `integer` data types) and categorical values (`factor`, `character`, and `logical`). The method also works on `matrix`, `list`, and `data.frame` data types by applying `smd()` over the columns of the `matrix` or `data.frame` and each item of the `list`.  The package is based on @yang2012unified.

```{r}
library(SugarMaryDenver)
```

# Examples

```{r numeric}
set.seed(123)
xn <- rnorm(100)
gg <- rep(c("A", "B"), each = 50)
smd(x = xn, g = gg)
```

```{r integer}
xi <- sample(1:20, 100, replace = TRUE)
smd(x = xi, g = gg)
```

```{r character}
xc <- unlist(replicate(2, sort(sample(letters[1:3], 50, replace = TRUE)), simplify = FALSE))
smd(x = xc, g = gg)
```

```{r factor}
xf <- factor(xc)
smd(x = xf, g = gg)
```

```{r logical}
xl <- as.logical(rbinom(100, 1, prob = 0.5))
smd(x = xl, g = gg)
```

```{r matrix}
mm <- cbind(xl, xl, xl, xl)
smd(x = mm, g = gg)
```

```{r list}
ll <- list(xn, xi, xf, xl)
smd(x = ll, g = gg)
```

```{r data.frame}
df <- data.frame(xn, xi, xc, xf, xl)
smd(x = df, g = gg)
```

## Using `smd` with `dplyr`

```{r dplyr}
library(dplyr)
df$g <- gg
df %>%
  summarize_at(
    .vars = vars(matches("^x")),
    .funs = funs(smd(., g = g)))
```


# Other standardized mean difference packages

See [tableone](https://CRAN.R-project.org/package=tableone) and [stddiff](https://cran.r-project.org/package=stddiff).

# References
